Sperm should evolve to make female meiosis fair.

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex

\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{}
\rhead{}
\renewcommand{\headrulewidth}{0.0pt}
\rfoot{}
\cfoot{\thepage}
\date{}
\bibliographystyle{plain}
\begin{document}
\maketitle
\begin{center}
$^1$Department of Plant Biology, \\ University of Minnesota – Twin
Cities, MN, 55108 
$^2$Center for Population Biology \& Department of Evolution and Ecology \\ University of California - Davis \\ Davis, CA, 95616
\end{center}

%FIgures 
{\bf Abstract:}
Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
Because only one of four products of female meiosis is transmitted to
offspring, o\"{o}genesis is inherently susceptible to such conflicts,
as alleles can compete to be transmitted into the egg. 
%(meiotic drive).  
Genomics conflicts drive the rapid evolution of suppression mechanisms, and
	many facets of  o\"{o}genesis and gametogenesis have been 
	interpreted as mechanisms of protection against genomic outlaws. 
Why then is female meiosis often left uncompleted until after fertilization - 
	potentially providing an opportunity for sperm alleles to
        meddle with its outcome in their own favor? %and facilitate meiotic drive? 
The population genetic theory presented herein suggests that
	sperm nearly always evolve to increase the fairness of female
        meiosis in the face of genomic conflicts. %%felt this
                                %%might be an important addition as
                                %%they will spread if no cost to organism
These results are consistent with current knowledge of sperm-dependent meiotic drivers 
(loci whose distortion of female meiosis depends on sperm genotype),
	and suggest that the requirement of fertilization for the
        completion of female meiosis, and variation in stage of meiosis that
        fertilization occurs at may represent a mechanism employed by females 
	 to ensure a fair meiosis. %prevent meiotic drive.
\newpage

>>>>

\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{}
\rhead{}
\renewcommand{\headrulewidth}{0.0pt}
\rfoot{}
\cfoot{\thepage}
\date{}
\bibliographystyle{plain}
\begin{document}
\maketitle
\begin{center}
$^1$Department of Plant Biology, \\ University of Minnesota --€“ Twin
Cities. St. Paul MN, 55108 \\
$^2$Center for Population Biology \& Department of Evolution and Ecology \\ University of California -- Davis. Davis, CA, 95616
\end{center}

%FIgures 
{\bf Abstract:}
Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
%Because only one of four products of female meiosis is transmitted to offspring, 
O\"{o}genesis is inherently susceptible to such conflicts
because alleles compete to be the  product of female meiosis transmitted to the egg. 
%(meiotic drive).  
Alleles that distort meiosis in their favor (i.e. meiotic drivers) often decrease organismal fitness,
	and therefore indirectly favor the evolution of mechanisms to suppress meiotic drive. % drive the rapid evolution of suppression mechanisms, and
In this light, many facets of  o\"{o}genesis and gametogenesis have been 
	interpreted as mechanisms of protection against genomic outlaws. 
Why then is female meiosis often left uncompleted until after fertilization - 
	potentially providing an opportunity for sperm alleles to
        meddle with its outcome and help like-alleles drive in heterozygous females? %and facilitate meiotic drive? 
The population genetic theory presented herein suggests that
	sperm nearly always evolve to increase the fairness of female
        meiosis in the face of genomic conflicts. %%felt this
                                %%might be an important addition as
                                %%they will spread if no cost to organism
These results are consistent with current knowledge of sperm-dependent meiotic drivers 
(loci whose distortion of female meiosis depends on sperm genotype),
	and suggest that the requirement of fertilization for the
        completion of female meiosis
        %, and variation in stage of meiosis that
        %fertilization occurs at 
        potentially represents a mechanism employed by females 
	 to ensure a fair meiosis. %prevent meiotic drive.
\newpage
